### 2023-03-16 chapters/paper_review/01_introduction.tex
Change reference keys for review paper (WIP)

%Hung-yi Lee: I think maybe we need a basic frame image. I happen to have an image that can be used here.
\begin{figure}
    \centering
    \includegraphics[width=0.80\textwidth]{paper_review/SSL_framework.png}
	 \caption{Framework for using self-supervised representation learning in
	 downstream applications}
    \label{fig:SSL_framework}
\end{figure}

% {\color{blue} Reviewer: Daniel, KL, SW}\\
% {\color{blue} Abdo -- this section is a work-in-progress}\\

% {\color{red} Hung-yi: Last time we have decided to use "pre-train" (with "-")} -- Done\\

% {\color{red} Hung-yi: Fine-tune? or finetune?} --Fine-tune Done\\

% {\color{red} Hung-yi: Fig. 1 or Figure 1} -- Fig. 1  Done\\

% \abdo{comment from Shinji: bring references to intro.}

% \kl{Comments from Nov. 10 meeting: 
% \begin{itemize}
%     \item \abdo{added.} would be good to define representation learning a bit more thoroughly in the intro, and mention that we are considering both discrete and continuous representations.
%     \item \abdo{added.} it would also be good to mention and broadly define "self-supervised" in the intro (as of now it's first mentioned in Section~\ref{sec:thirdwave}), and mention that the term is now used (and we use it) for some techniques that in the past would have just been called "unsupervised".
% \end{itemize}}

% \kl{New comments: 
% \begin{itemize}
%     \item It would be good to include somewhere in the paper a summary of tasks for which SSL pre-trained models have the SOTA performance.  Maybe it is in the paper already but I don't see it.
% \end{itemize}}

Over the past decade, deep learning approaches have revolutionized speech processing
through a giant leap in performance, enabling various real-world applications.
Supervised learning of deep neural networks has been the cornerstone of this
transformation, offering impressive gains for scenarios rich in labeled
data~\cite{lecun2015deep, hinton2012deep, bourlard2012connectionist}. 
Paradoxically, this heavy reliance on supervised learning has restricted progress in
languages and domains that do not attract the same level of labeling
investment. 
%Paradoxically, the heavy reliance on supervised learning bottlenecked the development of novel industrial speech applications and restricted progress in languages and domains that do not entertain the same level of labeling investment.

To overcome the need for labeled data, researchers have explored approaches that use
unpaired audio-only data to open up new industrial speech use-cases and
low-resource languages~\cite{Kemp1999, csl01_limsi, ma_bbn_06}. Inspired by how
children learn their first language through listening and interacting with
family and surroundings, scientists seek to use raw waveforms and
spectral signals to learn speech representations that capture low-level
acoustic events, lexical knowledge, all the way to syntactic and semantic
information. These learned representations are then used for target downstream
applications requiring a minimal number of labeled data~\cite{Hinton_2007,
LeCun06atutorial, bengio_representation_2013}. 
Formally, representation learning refers to algorithms for extracting latent
features that capture the underlying explanatory factors for the observed
input~\cite{bengio_representation_2013}. 
% Previous research efforts explored latent representations that are binary \cite{}, discrete \cite{} and continuous \cite{}. 
% These latent representations are generally used as inputs to a supervised predictor or a generation module. The whole network generating these representations is frequently used to initialize a supervised classifier~\cite{hinton_2006}.

%Learning representations approaches generally fall within \textit{unsupervised learning}, which refers to the family of machine learning methods that discover naturally occurring patterns in the training samples without any pre-assigned labels or scores associated with these samples~\cite{jordan_2015}. 
Representation learning approaches are generally considered examples of \textit{unsupervised
learning}, which refers to the family of machine learning methods that discover
naturally occurring patterns in training samples for which there are no pre-assigned
labels or scores~\cite{jordan_2015}. 
The term ``unsupervised'' is used to distinguish this family of methods from
``supervised'' approaches, which assign a label to each training sample, and
``semi-supervised'' approaches, which utilize a small number of training samples
with labels to guide learning using a larger volume of unlabeled samples.
Examples of unsupervised learning techniques include k-means clustering~\cite{vq}, mixture models~\cite{MoE}, autoencoders~\cite{hinton_94},
and non-negative matrix factorization~\cite{nmf}. 
\textit{Self-supervised learning} (SSL) is a fast-growing subcategory of
unsupervised learning approaches, which are techniques that utilize
information extracted from the input data itself as the label to learn
representations useful for 
% one or many 
downstream tasks. {\color{black} For example, unsupervised k-means clustering doesn't adhere to this definition of self-supervision since it iteratively minimizes the within-cluster variance during learning.}
In this review, we
focus on self-supervised learning approaches.

\Cref{fig:SSL_framework} outlines self-supervised representation learning in
relation to downstream applications. 
There are two stages in this framework.
In the first stage, we use SSL to pre-train a \textit{representation model},
also called an \textit{upstream model} or a \textit{foundation model}.
In the second stage, downstream tasks use either the learned
representation from the frozen model, or fine-tune the entire pre-trained model
in a supervised phase~\cite{hinton_2006}. 
%e.g., Automatic Speech Recognition (ASR) and Speaker Identification (SID) are two example downstream applications in \cref{fig:SSL_framework}. 
Automatic speech recognition (ASR) and speaker identification (SID) are 
examples of downstream applications in \cref{fig:SSL_framework}.
% There are some task-specific labeled data for each downstream task, and there is a downstream model on top of the upstream model. The labeled data is used to train the corresponding downstream models and optionally fine-tune upstream models.
% If the pre-training stage only leverages unlabeled data, the framework in \cref{fig:SSL_framework} is named self-supervised learning (SSL), which is the focus of this review paper.
% Pre-training with labeled data is out of the scope. Before the term self-supervised learning was widely used, pre-training without any labeled data was usually known as an \texit{unsupervised learning} approach.

%Lee: in the introduction, we may need a specific example. Below is just an example.
%With unsupervised ASR, imagine that in the future, when an indigenous language family buys an intelligent assistant, even if the assistant does not support the language at the beginning, it would automatically learn to transcribe the new language after learning.

% The speech representation learning approaches evolved over the past few decades through at least three major stages, as discussed in \cref{sec:thirdwave}; static and temporal clustering algorithms \cite{}, energy-based neural models \cite{}, and more recently, self-supervised learning techniques \cite{}. 
% Across all these three waves of representation learning, three main desirable characteristics emerged; features need to be disentangled, invariant, and hierarchical. 
It is considered desirable for learned speech representations to be
disentangled, invariant, and hierarchical.
Since spoken utterances contain much richer information than the corresponding text
transcriptions---e.g., speaker identity, style, emotion, surrounding noise, and
communication channel noise---it is important to learn representations that
disentangle these factors of variation. Furthermore, invariance of the learned
features to changes in background noise and in the communication channel ensures
stability with respect to downstream application scenarios. Learning feature
hierarchies at the acoustic, lexical, and semantic levels supports applications
with different requirements. For instance, whereas a speaker identification task
benefits from a low-level acoustic representation, a speech
translation task requires a more semantic representation of the input
utterance.

Due to the popularity of SSL, reviews have been published about the
technology in general~\cite{bommasani2021opportunities,ericsson2021selfsupervised,LiuSSLsurvey} as well as its application to natural language processing (NLP)~\cite{rogers-etal-2020-primer,liu2021pretrain,xia-etal-2020-bert,QiuSSLNLPsurvey} and computer vision (CV)~\cite{JingSSLCVsurvey}. Recently, a brief overview with a general focus on speech representation learning was published \cite{borgholt_22}. However, none of these overviews focus exclusively on SSL for speech processing. Since the speech signal differs greatly from image and text inputs, many  theories  and technologies have been developed to address the unique challenges of speech. One review addresses speech representation learning based on deep learning models~\cite{latif2021deep}, but does not address recent developments in self-supervised learning. This motivates this overview of speech SSL.

%For survey papers on TL to speech processing, please check the reference~\cite{survey_speech_TL}.
%The representation learning methods covered in this review are primarily task-agnostic methods that only benefit from supervised data at the fine-tuning stage, as shown in \cref{}. %Lee: Similar sentence may have been mentioned in the paragraph describing the figure.

>>>>

\begin{figure}
    \centering
    \includegraphics[width=0.80\textwidth]{paper_review/SSL_framework.png}
	 \caption{Framework for using self-supervised representation learning in
	 downstream applications}
    \label{fig:SSL_framework}
\end{figure}

Over the past decade, deep learning approaches have revolutionized speech processing
through a giant leap in performance, enabling various real-world applications.
Supervised learning of deep neural networks has been the cornerstone of this
transformation, offering impressive gains for scenarios rich in labeled
% data~\cite{lecun_deep_2015}. 
% data~\cite{lecun_deep_2015, hinton_deep_2012}. 
% data~\cite{bourlard_connectionist_1994}. 
% data~\cite{lecun_deep_2015,bourlard_connectionist_1994}. 
data~\cite{lecun_deep_2015,hinton_deep_2012,bourlard_connectionist_1994}. 
Paradoxically, this heavy reliance on supervised learning has restricted progress in
languages and domains that do not attract the same level of labeling
investment.

To overcome the need for labeled data, researchers have explored approaches that use
unpaired audio-only data to open up new industrial speech use-cases and
low-resource languages~\cite{kemp_unsupervised_1999, lamel_lightly_2002, ma_unsupervised_2006}. Inspired by how
children learn their first language through listening and interacting with
family and surroundings, scientists seek to use raw waveforms and
spectral signals to learn speech representations that capture low-level
acoustic events, lexical knowledge, all the way to syntactic and semantic
information. These learned representations are then used for target downstream
applications requiring a minimal number of labeled data~\cite{hinton_learning_2007,
lecun_tutorial_2006, bengio_representation_2013}. 
Formally, representation learning refers to algorithms for extracting latent
features that capture the underlying explanatory factors for the observed
input~\cite{bengio_representation_2013}.

Representation learning approaches are generally considered examples of \textit{unsupervised
learning}, which refers to the family of machine learning methods that discover
naturally occurring patterns in training samples for which there are no pre-assigned
labels or scores~\cite{jordan_machine_2015}. 
The term ``unsupervised'' is used to distinguish this family of methods from
``supervised'' approaches, which assign a label to each training sample, and
``semi-supervised'' approaches, which utilize a small number of training samples
with labels to guide learning using a larger volume of unlabeled samples.
Examples of unsupervised learning techniques include k-means clustering~\cite{gray_vector_1984}, mixture models~\cite{jordan_hierarchical_1994}, autoencoders~\cite{hinton_autoencoders_1994},
and non-negative matrix factorization~\cite{lee_learning_1999}. 
\textit{Self-supervised learning} (SSL) is a fast-growing subcategory of
unsupervised learning approaches, which are techniques that utilize
information extracted from the input data itself as the label to learn
representations useful for 
downstream tasks. {\color{black} For example, unsupervised k-means clustering doesn't adhere to this definition of self-supervision since it iteratively minimizes the within-cluster variance during learning.}
In this review, we
focus on self-supervised learning approaches.

\Cref{fig:SSL_framework} outlines self-supervised representation learning in
relation to downstream applications. 
There are two stages in this framework.
In the first stage, we use SSL to pre-train a \textit{representation model},
also called an \textit{upstream model} or a \textit{foundation model}.
In the second stage, downstream tasks use either the learned
representation from the frozen model, or fine-tune the entire pre-trained model
in a supervised phase~\cite{hinton_reducing_2006}. 
Automatic speech recognition (ASR) and speaker identification (SID) are 
examples of downstream applications in \cref{fig:SSL_framework}.

It is considered desirable for learned speech representations to be
disentangled, invariant, and hierarchical.
Since spoken utterances contain much richer information than the corresponding text
transcriptions---e.g., speaker identity, style, emotion, surrounding noise, and
communication channel noise---it is important to learn representations that
disentangle these factors of variation. Furthermore, invariance of the learned
features to changes in background noise and in the communication channel ensures
stability with respect to downstream application scenarios. Learning feature
hierarchies at the acoustic, lexical, and semantic levels supports applications
with different requirements. For instance, whereas a speaker identification task
benefits from a low-level acoustic representation, a speech
translation task requires a more semantic representation of the input
utterance.

Due to the popularity of SSL, reviews have been published about the
technology in general~\cite{bommasani_opportunities_2021,ericsson_selfsupervised_2022,liu_selfsupervised_2021} as well as its application to natural language processing (NLP)~\cite{rogers_primer_2020,liu_pretrain_2021,xia_which_2020,qiu_pretrained_2020} and computer vision (CV)~\cite{jing_selfsupervised_2021}. Recently, a brief overview with a general focus on speech representation learning was published \cite{borgholt_brief_2022}. 
However, none of these overviews focus exclusively on SSL for speech processing. Since the speech signal differs greatly from image and text inputs, many  theories  and technologies have been developed to address the unique challenges of speech. 
One review addresses speech representation learning based on deep learning models~\cite{latif_deep_2021}, but does not address recent developments in self-supervised learning. This motivates this overview of speech SSL.

### 2023-03-16 chapters/paper_review/02_three_waves.tex
Change reference keys for review paper (WIP)

% {\color{blue} Reviewer: KL, KK, Daniel }\\
In this section we present the historical background of the current surge in
self-supervised representation learning methods in the context of two previous
waves of research work in the 1990s and 2000s. The discussed approaches go
beyond speech to describe the overall landscape of machine learning development during the
past few decades.

%%KK: if you adopt the paragraph below I suggest renaming this as "Features based on generative models" or something
\subsection{Clustering and mixture models}
% {\color{blue} Tara}\\

%\kl{to people who are familiar with modern rep learning work, this first wave may not seem like rep learning at all.  readers might also wonder why we are talking about what look like standard supervised ASR in the 3rd paragraph below.  one option is to describe it is as unsupervised learning of latent variables (Gaussian components, HMM states) within a supervised model. (something like that -- I haven't thought about the wording carefully)}
%KK: suggesting the paragraph below for inclusion here
%The initial wave of speech representation {\em learning} focused on extracting speech features from generative models. The main objective was
%to render  the knowledgpe learned by generative models accessible to
%discriminative downstream classifiers, or to map variable-length sequences into  fixed-length representations.
%The generative model, typically a global Gaussian mixture model (GMM) or a Hidden Markov Model (HMM) was trained by
%by maximizing the likelihood of the data given the
%model, which could be done in either an unsupervised or a supervised way. Feature vectors were then derived from the parameters of the trained model. In
%the case of {\em Fisher vectors}, the features were the normalized gradients of the log-likelihood
%with respect to the model parameters (mixture weights, means, variances) of the Gaussian mixtures.
%An extension of this approach (likelihood ratio score space) used the derivative of the log likelihood ratio of	two models,
%e.g., a background model and a foreground model.
%Examples of their use in speech processing include speech recognition \cite{smith01,venkata03} and speaker recognition \cite{wan03}.
%Subsequent techniques in speaker and language verification \cite{dehak11a,dehak11b} similarly extracted parameters (concatenated means) from trained background GMMs as representations that were then %combined with low-rank projections of speaker/session or language specific vectors.

@@

Early work used simple clustering methods. For example, in work such
as \cite{Rabiner1979,Wilpon1985}, word patterns were clustered
semi-automatically using techniques such as k-means, after which isolated words
were recognized by finding the training cluster closest to 
  the test data.   % AMH: check

Through time, modeling techniques improved such that subword units were
represented by Gaussian mixture models (GMMs) \cite{Gauvain1994}, which facilitated
the modeling of more variability in the input data. GMMs were first built for
context-independent phonemes; state-clustering 
algorithms~\cite{Young1994} then resulted in GMMs for context-dependent phonemes. 
Each latent component of these mixture models acted as a template of a
prototypical speech frame, 
% leading to their deficiency to deal with 
  making it difficult to handle             % AMH: check
large volumes of data with diverse characteristics. 
Furthermore, dynamical models like hidden Markov models (HMMs)~\cite{Bahl1986}
allowed for the processing of continuous speech rather than just isolated word
recognition. These generative GMM and HMM models were trained by maximizing the
likelihood of data given the model, which could be accomplished in either an
unsupervised or a supervised manner.

Another line of research focused on extracting speech features from generative
models. The main objective here was to render the knowledge learned by generative
models accessible to discriminative downstream classifiers, or to map
variable-length sequences to fixed-length representations. Feature vectors
were derived from the parameters of trained GMM models. In the case of {\em
Fisher vectors}, the features were the normalized gradients of the
log-likelihood with respect to the model parameters (mixture weights, means, and
variances) of the Gaussian mixtures. An extension of this approach (likelihood
ratio score space) used the derivative of the log-likelihood ratio of two
models, e.g., a background model and a foreground model. Examples of their use
in speech processing include speech recognition~\cite{smith01,venkata03} and
speaker recognition~\cite{wan03}. Subsequent techniques in speaker and language
verification~\cite{dehak11a,dehak11b} similarly extracted parameters
(concatenated means) from trained background GMMs as representations that were
then combined with low-rank projections of speaker/session- or language-specific
vectors.

@@

%{\color{blue} Tara}\\
More recently, representation learning has seen a shift of focus towards neural models,
which, compared to GMMs and HMMs, offer distributed representations with more
capacity to model diverse input signals into efficient latent binary codes. 
Examples of early techniques include restricted Boltzmann machines
(RBM)~\cite{hinton_2006}, denoising autoencoders~\cite{DAE}, noise contrastive
estimation (NCE)~\cite{gutmann2012noise}, sparse coding~\cite{Olshausen1996,
sparse_lee, sivaram2010sparse}, and energy-based methods~\cite{Ranzato2007}.
Many of these techniques have also been applied to CV and NLP problems, which
provided inspiration for their application to speech.
%Examples of techniques include Restricted Boltzmann Machine (RBM) \cite{hinton2012practical,Srivastava2013}, Neural Auto Encoders \cite{Lecun2012}, Noise Contrastive Estimation (NCE) \cite{chen2015recurrent},  Sparse Coding \cite{sivaram2010sparse} and Energy-based method \cite{Ranzato2007}. Note that many of these techniques were also tried for CV and NLP problems, which provided inspiration to try them for ASR as well.

Higher-capacity neural models were achieved by stacking several neural network
layers to build progressively higher-level concept representations.
However, these deeper networks also increased the training complexities. For
example, approximate training methods such as contrastive divergence
\cite{hinton_cd_2002} were a practical technique to 
% make       RBM training more efficient. 
  streamline RBM training.                   % AMH: check
Furthermore, deep networks had non-convex objective functions, which
often resulted in long training times compared to GMMs, which are trained
using full batches instead of mini-batch learning.

% http://proceedings.mlr.press/v2/ranzato07a/ranzato07a.pdf, 
%https://auai.org/uai2013/prints/papers/166.pdf, % https://www.cs.cmu.edu/~rsalakhu/papers/neco_DBM.pdf
% problems with this trend: approximate gradient, failure modes, long training time, no clear wins in downstream applications.

@@

% {\color{blue} Hung-yi + Abdo}\\

%Note on 12/09: We will mainly describe the speech part but link to CV and NLP. 
%discuss instance classification, contrastive losses,\\

%Currently, the network learns a function that maps input to desired representations by solving a \textit{pretext} task.
%After pre-training, learned representation models could be applied to downstream tasks through feature-based representation extraction or fine-tuning as part of the downstream model.  
%By optimizing carefully designed pretext tasks, neural models can capture latent representations which generalize well across a wide range of downstream applications, usually under few-shot or even zero-shot conditions. 
A more recent trend is learning networks that map the input to desired
representations by solving a \textit{pretext} task. Such studies have several
characteristics:
(1)~All layers are trained end-to-end to optimize a single pretext task instead
of relying on layer-wise pre-training
(2)~Past stacked networks typically had only a few layers, but 
very deep networks with more than ten layers are now common.
(3)~It is common to evaluate a representation model on a wide range of tasks.
For example, in NLP, a representation model is usually assessed on GLUE,
which comprises nine tasks~\cite{GLUE}, whereas in speech, a representation model can be
evaluated on SUPERB, which comprises ten tasks~\cite{yang21c_interspeech}, 
as described in detail in \cref{sec:benchmark}.

The cornerstone of this third wave is the design of a pretext task, which
allows the model to efficiently leverage knowledge from unlabeled data.
The pretext task should be challenging enough for the model to learn high-level
abstract representations and 
% not be too amiable to exploit low-level shortcuts.
  not be so easy as to encourage the exploitation of low-level shortcuts.  % AMH: check
Early breakthroughs included end-to-end learning of deep neural architectures
via pretext tasks for restoring the true color of black-and-white
images~\cite{colorizing}, joint learning of latent representations and their
cluster assignments~\cite{deepcluster}, and the prediction of the relative positions of
image patches~\cite{context_pred}. Other popular approaches include variational
autoencoders (VAEs)~\cite{vae, rezende2014stochastic}. While typical autoencoders learn data
representations using unsupervised objectives by reconstructing the input
after passing it through an information bottleneck, VAEs estimate a neural model of a probability density function (pdf) that approximates the unknown “true” distribution of the observed data, for which we only have access to independently identically distributed (iid) samples. It is also important to mention dynamical VAEs \cite{Girin2021}, which is an extension of VAE for sequential data such as speech.

In the SSL context, a pretext task related to autoencoding is to generate
an object from its partial information. Such tasks are widely used in NLP, for
example, using 
the previous tokens 
in a sentence to predict the next token such as in
ELMo~\cite{peters2018deep}, the GPT
series~\cite{alex2019GPT2}, and %brown2020gpt3 alex2018GPT,
Megatron~\cite{shoeybi2020megatronlm}, or predicting the masked tokens in a
sentence such as with the bidirectional encoder representations from Transformers
(BERT) series~\cite{jacob2019BERT,Liu2019RoBERTa}. %,albert
%ELMo~\cite{peters2018deep}, GPT series~\cite{alex2018GPT,alex2019GPT2,brown2020gpt3}, Megatron~\cite{shoeybi2020megatronlm} are gigantic LM which predict the next token. 
%BERT~\cite{jacob2019BERT}, RoBERTa~\cite{Liu2019RoBERTa}, ALBERT~\cite{albert} masked the input token, and the models are learned to predict the masked tokens. 
Another common pretext task in the third wave is contrastive
learning~\cite{oord2018representation}, in which a model learns to identify a
target instance from a set of negative samples. 
This approach has become especially popular in the
CV context~\cite{pmlr-v119-chen20j,he2020momentum,chen2020improved,SwAV}.
In this survey, we will mainly focus on techniques for pretext task
optimization for speech processing, and discuss these techniques in detail
in \cref{sec:approach}.

\subsection{Other related work}
A closely related area of research that is not covered in this review is
semi-supervised pre-training methods such as pseudo-labeling (that is,
self-training). Pseudo-labeling (PL) relies on a supervised teacher model to
label a large volume of speech-only data, which is then used to augment the
initial labeled data to train a student model~\cite{Kemp1999, csl01_limsi,
ma_bbn_06, hari_1mhour}. PL has been successful and widely adopted in the
speech community since the 1990s. Other proposed variations of PL include
augmenting speech-only data with noise to improve robustness, iterating
over the PL process to improve teacher labeling quality, and training student
models with more parameters than their original teachers to capture the
complexities in vastly larger speech-only data~\cite{park2020improved,
xu2020iterative, xiao_scaling_2021}. 
Both SSL and PL leverage unlabeled speech-only data.
One distinguishing factor in PL is the utilization of supervised data for a
specific task during model pre-training, which limits the model's focus to
a single (or at best a few) downstream tasks. 
SSL, in turn, is an attempt to learn task-agnostic representations to benefit
a wide range of tasks.

Transfer learning (TL) is another closely related area of research for
pre-training speech models. TL transfers knowledge captured by models
trained on one task to different but related tasks~\cite{caruana1997multitask}. 
The past few decades have seen active research on TL and its
extension to multitask learning for more general representations. 
Multilingual and cross-lingual supervised models have proven superior in
low-resource speech recognition tasks~\cite{Cui2015MultilingualRF}.
SSL can be regarded as a type of TL because knowledge learned from pre-training
is used for different downstream tasks.
This survey paper focuses on SSL, and not all TL technologies for speech. 
One survey indeed addresses TL for speech processing~\cite{survey_speech_TL}
but does not include current SSL technologies for speech.

%Then contrastive loss is applied~\cite{oord2018representation}, in which the model learns to classify the positive instances from a set of unrelated negative samples. SimCLR~\cite{pmlr-v119-chen20j}, MoCo~\cite{he2020momentum,chen2020improved} and SwAV~\cite{SwAV} are representative ones. 
%These approaches learn representations from visual inputs by maximizing agreement between differently augmented views of the same sample while minimizing the agreement between different samples. 
%Then there is a series of work trying to remove the requirement of negative examples. 
%BYOL~\cite{BYOL} trains a student network to predict the representations of the teacher network.  The parameter of the teacher network is the exponential moving average of the previous student network. In such a scenario, negative examples are not required. Then SimSiam~\cite{Chen_2021_CVPR} improved BYOL by removing the moving average and found that good representation can still be learned in this way.
%In Barlow twins~\cite{BarlowTwins}, the cross-correlation matrix between the representations of two identical encoders fed with distorted versions of a sample is measured. The encoder learns to make the cross-correlation matrix as close to the identity matrix as possible.
%VICReg~\cite{bardes2021vicreg}

%The 3rd wave begins from NLP and CV communities. %Is it true?
%Below the progress of self-supervised learning in CV  and NLP are briefly summarized. %(However, we have to consider how to connect this part with other parts.)

%In CV, some earlier approaches make some pretexts to train instance classifiers, and the last hidden layers of the classifiers are used as the representations in the downstream tasks. Below are some representative approaches:
%\begin{itemize}
%\item Position~\cite{Doersch_2015_ICCV}: predicting the relative position between two random patches from one image. A model needs to understand the spatial context of objects to tell the relative position between parts.
%\item Jigsaw~\cite{noroozi2016unsupervised}: The model is trained to place shuffled patches back to the original locations.
%\item Rotation~\cite{gidaris2018unsupervised}: The model is trained to predict which rotation has been applied.
%\end{itemize}

%How to deal with negative examples is critical in this series. SimCLR~\cite{pmlr-v119-chen20j} uses a large batch size to obtain a lot of negative examples, while MoCo uses a momentum encoder to make the negative samples’ representation consistent.

%In ELECTRA~\cite{clark2020electra}, given a sentence with some tokens replaced, the ELECTRA model predicts whether a token is generated or not. 
%XLNet~\cite{XLNet}
%MASS~\cite{MASS}, BART~\cite{lewis-etal-2020-bart}, and T5~\cite{T5} are seq2seq models. The input is disturbed by different approaches, and the seq2seq model learns to reconstruct the disturbed input. %(There are speech T5, should we mention this work?)
%The contrastive approaches have been applied at the sentence level in NLP~\cite{gao-etal-2021-simcse,kim-etal-2021-self,yan-etal-2021-consert,giorgi-etal-2021-declutr}.

%Turing NLG only has webpage: https://www.microsoft.com/en-us/research/blog/turing-nlg-a-17-billion-parameter-language-model-by-microsoft/
%I don't know how to cite it.

%ref: http://speech.ee.ntu.edu.tw/~tlkagk/courses/DLHLP20/BERT%20train%20(v8).pdf

>>>>

In this section we present the historical background of the current surge in
self-supervised representation learning methods in the context of two previous
waves of research work in the 1990s and 2000s. The discussed approaches go
beyond speech to describe the overall landscape of machine learning development during the
past few decades.

\subsection{Clustering and mixture models}

@@

Early work used simple clustering methods. For example, in work such
as \cite{rabiner_considerations_1979,wilpon_modified_1985}, word patterns were clustered
semi-automatically using techniques such as k-means, after which isolated words
were recognized by finding the training cluster closest to 
  the test data.

Through time, modeling techniques improved such that subword units were
represented by Gaussian mixture models (GMMs) \cite{gauvain_maximum_1994}, which facilitated
the modeling of more variability in the input data. GMMs were first built for
context-independent phonemes; state-clustering 
algorithms~\cite{young_state_1994} then resulted in GMMs for context-dependent phonemes. 
Each latent component of these mixture models acted as a template of a
prototypical speech frame, 
making it difficult to handle 
large volumes of data with diverse characteristics. 
Furthermore, dynamical models like hidden Markov models (HMMs)~\cite{bell_adaptation_2021}
allowed for the processing of continuous speech rather than just isolated word
recognition. These generative GMM and HMM models were trained by maximizing the
likelihood of data given the model, which could be accomplished in either an
unsupervised or a supervised manner.

Another line of research focused on extracting speech features from generative
models. The main objective here was to render the knowledge learned by generative
models accessible to discriminative downstream classifiers, or to map
variable-length sequences to fixed-length representations. Feature vectors
were derived from the parameters of trained GMM models. In the case of {\em
Fisher vectors}, the features were the normalized gradients of the
log-likelihood with respect to the model parameters (mixture weights, means, and
variances) of the Gaussian mixtures. An extension of this approach (likelihood
ratio score space) used the derivative of the log-likelihood ratio of two
models, e.g., a background model and a foreground model. Examples of their use
in speech processing include speech recognition~\cite{smith_speech_2001,venkataramani_support_2003} and
speaker recognition~\cite{wan_svmsvm_2003}. Subsequent techniques in speaker and language
verification~\cite{dehak_frontend_2011,dehak_language_2011} similarly extracted parameters
(concatenated means) from trained background GMMs as representations that were
then combined with low-rank projections of speaker/session- or language-specific
vectors.

@@

More recently, representation learning has seen a shift of focus towards neural models,
which, compared to GMMs and HMMs, offer distributed representations with more
capacity to model diverse input signals into efficient latent binary codes. 
Examples of early techniques include restricted Boltzmann machines
(RBM)~\cite{hinton_reducing_2006}, denoising autoencoders~\cite{vincent_stacked_2010}, noise contrastive
estimation (NCE)~\cite{gutmann_noisecontrastive_2012}, sparse coding~\cite{olshausen_emergence_1996,
lee_efficient_2006, sivaram_sparse_2010}, and energy-based methods~\cite{ranzato_unified_2007}.
Many of these techniques have also been applied to CV and NLP problems, which
provided inspiration for their application to speech.

Higher-capacity neural models were achieved by stacking several neural network
layers to build progressively higher-level concept representations.
However, these deeper networks also increased the training complexities. For
example, approximate training methods such as contrastive divergence
\cite{hinton_training_2002} were a practical technique to streamline RBM training. 
Furthermore, deep networks had non-convex objective functions, which
often resulted in long training times compared to GMMs, which are trained
using full batches instead of mini-batch learning.

@@

A more recent trend is learning networks that map the input to desired
representations by solving a \textit{pretext} task. Such studies have several
characteristics:
(1)~All layers are trained end-to-end to optimize a single pretext task instead
of relying on layer-wise pre-training
(2)~Past stacked networks typically had only a few layers, but 
very deep networks with more than ten layers are now common.
(3)~It is common to evaluate a representation model on a wide range of tasks.
For example, in NLP, a representation model is usually assessed on GLUE,
which comprises nine tasks~\cite{wang_glue_2018}, whereas in speech, a representation model can be
evaluated on SUPERB, which comprises ten tasks~\cite{yang_superb_2021}, 
as described in detail in \cref{sec:benchmark}.

The cornerstone of this third wave is the design of a pretext task, which
allows the model to efficiently leverage knowledge from unlabeled data.
The pretext task should be challenging enough for the model to learn high-level
abstract representations and 
% not be too amiable to exploit low-level shortcuts.
  not be so easy as to encourage the exploitation of low-level shortcuts.  % AMH: check
Early breakthroughs included end-to-end learning of deep neural architectures
via pretext tasks for restoring the true color of black-and-white
images~\cite{zhang_colorful_2016}, joint learning of latent representations and their
cluster assignments~\cite{caron_deep_2018}, and the prediction of the relative positions of
image patches~\cite{doersch_unsupervised_2016}. Other popular approaches include variational
autoencoders (VAEs)~\cite{kingma_autoencoding_2014, rezende_stochastic_2014}. While typical autoencoders learn data
representations using unsupervised objectives by reconstructing the input
after passing it through an information bottleneck, VAEs estimate a neural model of a probability density function (pdf) that approximates the unknown “true” distribution of the observed data, for which we only have access to independently identically distributed (iid) samples. It is also important to mention dynamical VAEs \cite{girin_dynamical_2021}, which is an extension of VAE for sequential data such as speech.

In the SSL context, a pretext task related to autoencoding is to generate
an object from its partial information. Such tasks are widely used in NLP, for
example, using 
the previous tokens 
in a sentence to predict the next token such as in
ELMo~\cite{peters_deep_2018}, the GPT
series~\cite{radford_language_2019}, and 
Megatron~\cite{shoeybi_megatronlm_2020}, or predicting the masked tokens in a
sentence such as with the bidirectional encoder representations from Transformers
(BERT) series~\cite{devlin_bert:_2018,liu_roberta_2019}. 
Another common pretext task in the third wave is contrastive
learning~\cite{oord_representation_2018}, in which a model learns to identify a
target instance from a set of negative samples. 
This approach has become especially popular in the
CV context~\cite{chen_simple_2020,he_momentum_2020,chen_improved_2020,caron_unsupervised_2020}.
In this survey, we will mainly focus on techniques for pretext task
optimization for speech processing, and discuss these techniques in detail
in \cref{sec:approach}.

\subsection{Other related work}
A closely related area of research that is not covered in this review is
semi-supervised pre-training methods such as pseudo-labeling (that is,
self-training). Pseudo-labeling (PL) relies on a supervised teacher model to
label a large volume of speech-only data, which is then used to augment the
initial labeled data to train a student model~\cite{kemp_unsupervised_1999, lamel_lightly_2002,
ma_unsupervised_2006, parthasarathi_lessons_2019}. PL has been successful and widely adopted in the
speech community since the 1990s. Other proposed variations of PL include
augmenting speech-only data with noise to improve robustness, iterating
over the PL process to improve teacher labeling quality, and training student
models with more parameters than their original teachers to capture the
complexities in vastly larger speech-only data~\cite{park_improved_2020,
xu_iterative_2020, xiao_scaling_2021}. 
Both SSL and PL leverage unlabeled speech-only data.
One distinguishing factor in PL is the utilization of supervised data for a
specific task during model pre-training, which limits the model's focus to
a single (or at best a few) downstream tasks. 
SSL, in turn, is an attempt to learn task-agnostic representations to benefit
a wide range of tasks.

Transfer learning (TL) is another closely related area of research for
pre-training speech models. TL transfers knowledge captured by models
trained on one task to different but related tasks~\cite{caruana_multitask_1997}. 
The past few decades have seen active research on TL and its
extension to multitask learning for more general representations. 
Multilingual and cross-lingual supervised models have proven superior in
low-resource speech recognition tasks~\cite{cui_multilingual_2015}.
SSL can be regarded as a type of TL because knowledge learned from pre-training
is used for different downstream tasks.
This survey paper focuses on SSL, and not all TL technologies for speech. 
One survey indeed addresses TL for speech processing~\cite{bell_adaptation_2021}
but does not include current SSL technologies for speech.
